Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-15 ICML-submission/from-weak-separability-to-strong-separability.tex
almost finished pass over section 5

In this section, we construct and analyze a mapping $\phi$ from the unit ball of
$\R^d$ into a infinite-dimensional inner product space that has the property
that if a set of labeled examples in the unit ball is \emph{weakly} linearly
separable with a margin $\gamma$ after applying the mapping $\phi$ the examples
become \emph{strongly} linearly separable with a margin $\gamma'$ and their
norms are bounded by $2$. The parameter $\gamma'$ is a function of the old
margin $\gamma$ and the number of classes $K$, and is specified in the
\autoref{theorem:margin-transformation} below.

The domain of $\phi$ is the unit ball of $\R^d$. The co-domain of $\phi$ is the
classical real separable Hilbert space $\ell_2 = \{ x \in \R^\infty ~:~
\sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped with the standard inner product
$\ip{x}{x'}_{\ell_2} = \sum_{i=1}^\infty x_i x'_i$. We index the coordinates of
$\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of non-negative
integers. The mapping is defined by
\begin{align*}
& \left(\phi(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \\
& \qquad \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\end{align*}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2! \dots \alpha_d!}$
is the multinomial coefficient. The corresponding kernel is
$$
k(x,x') = \ip{\phi(x)}{\phi(x')}_{\ell_2} = \frac{1}{1 - \frac{1}{2}\ip{x}{x'}_{\R^d}} \; .
$$
We note that the formula for $k$ implies that $k(x,x) < +\infty$ for any $x$ in
the unit ball of $\R^d$ and therefore $\phi(x)$ lies in $\ell_2$. We also note
that kernel function $k(x,x')$ can be evaluated in $O(d)$ time.

@@

The idea behind construction and analysis of the mapping $\phi$ is polynomial
approximation. According to the well known Stone-Weierstrass theorem (see
e.g.~\citet[Section~10.10]{Davidson-Donsig-2010}), on a compact set,
multivariate polynomials uniformly approximate any continuous function.
Intuitively speaking, we use a multivariate polynomial to approximate, on the
unit ball of $\R^d$, the indicator function corresponding to the intersection of
$m=K-1$ halfspaces. Within margin $\gamma$ along the decision boundary, we allow
the polynomial to attain arbitrary value. The polynomial separates examples in
one class from examples in the other classes. To be able to quantify $\gamma'$
we need to quantify certain parameters of the approximating polynomial. We
construct two different polynomials with different parameters. The parameters
are quantified in
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
stated below.

@@

\begin{proof}
First osberve that for any $t=1,2,\dots,T$,
$$
\norm{\phi(x_t)}_{\ell_2} = k(x_t,x_t) = \frac{1}{1 - \frac{1}{2} \norm{x_t}^2} \le 2 \; .
$$

>>>>

In this section, we focus on the setting of online bandit multiclass
linear prediction, where the set of examples shown by the adversary is weakly
linearly separable (Definition~\ref{definition:weak-linear-separability}).
Throughout this section, we assume without loss of generality that
all examples lie in $\B(0,1)$, the unit $\ell_2$ ball of
$\R^d$.\footnote{Consider the following preprocessing step: upon seeing every
example $x_t$, we define $\hat{x}_t = \frac{x_t}{\|x_t\|}$ as its $\ell_2$ normalization.
It can be easily checked that if $(x_1,y_1), \ldots, (x_T,y_T)$ is weakly linearly
separable by margin $\gamma$ and all $\| x_t \|$'s' have norm at most $R$, then the set of
normalized examples $(\hat{x}_1,y_1), \ldots, (\hat{x}_T,y_T)$ is weakly linearly
separable by margin $\gamma/R$.}
Note that Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
alone does not guarantee a finite mistake bound in this setting, as weak linear separability
does not imply strong linear separability.

%from the unit ball of
%$\R^d$ into a infinite-dimensional inner product space that

In this section, we consider a kernel function $k(\cdot, \cdot)$
whose corresponding feature map $\phi(\cdot)$ can transform a set of
\emph{weakly} linear separable examples to \emph{strongly} separable.
Specifically, $\phi$
has the property
that if a set of labeled examples in the unit ball is weakly linearly
separable with a margin $\gamma$, then after applying $\phi$ the examples
become strongly linearly separable with a margin $\gamma'$ and their
norms are bounded by $2$. The parameter $\gamma'$ is a function of the old
margin $\gamma$ and the number of classes $K$, and is specified in
\autoref{theorem:margin-transformation} below.

Specifically, the kernel function $k: \B(0,1) \times \B(0,1) \to \R$ is defined as:
$$
k(x,x') = \frac{1}{1 - \frac{1}{2}\ip{x}{x'}_{\R^d}} \; ,
$$
We note that the formula for $k$ implies that $k(x,x) < +\infty$ for any $x$ in
$\B(0,1)$. We also note
that kernel function $k(x,x')$ can be evaluated in $O(d)$ time.

Consider the
classical real separable Hilbert space $\ell_2 = \{ x \in \R^\infty ~:~
\sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped with the standard inner product
$\ip{x}{x'}_{\ell_2} = \sum_{i=1}^\infty x_i x'_i$.
Through indexing the coordinates of
$\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of non-negative
integers, the feature map that corresponds to $k$, $\phi: \B(0,1) \to \ell_2$, is defined as follows:
\begin{align*}
& \left(\phi(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \\
& \qquad \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\end{align*}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2! \dots \alpha_d!}$
is the multinomial coefficient.
It can be easily checked that
\[ k(x,x') = \ip{\phi(x)}{\phi(x')}_{\ell_2}. \]

%The mapping is defined by
%and therefore $\phi(x)$ lies in $\ell_2$.
%The domain of $\phi$ is the unit ball of $\R^d$. The co-domain of $\phi$ is

@@

\subsection{Proof of Theorem~\ref{theorem:margin-transformation}}
The idea behind construction and analysis of the mapping $\phi$ is polynomial
approximation. According to the well known Stone-Weierstrass theorem (see
e.g.~\citet[Section~10.10]{Davidson-Donsig-2010}), on a compact set,
multivariate polynomials uniformly approximate any continuous function.
Intuitively speaking, we use a multivariate polynomial to approximate, on the
unit ball of $\R^d$, the indicator function corresponding to the intersection of
$m=K-1$ halfspaces. Within margin $\gamma$ along the decision boundary, we allow
the polynomial to attain arbitrary value. The polynomial separates examples in
one class from examples in the other classes. To be able to quantify $\gamma'$
we need to quantify certain parameters of the approximating polynomial. We
construct two different polynomials with different parameters. The parameters
are quantified in
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
stated below.

@@

\begin{proof}[Proof of Theorem~\ref{theorem:margin-transformation}]
First osberve that for any $t=1,2,\dots,T$,
$$
\norm{\phi(x_t)}_{\ell_2} = k(x_t,x_t) = \frac{1}{1 - \frac{1}{2} \norm{x_t}^2} \le 2 \; .
$$
